Large scale risks from pesticides in small streams

### 2016-10-17 article.tex
Update on Overleaf.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% 150-200 words
Small streams are important refugia for biodiversity.
In agricultural areas they may be at high risk from pesticide pollution. 
However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales. 
We used data from German governmental water quality monitoring to quantify the drivers of pesticide risk and to assess pesticide risk in small streams on a large scale. 
The data set comprised of 2,918,604 measurements related to 42,236 samples in 3,049 sampling sites and of 484 pesticides. 
We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification. 
Agriculture land use lead to an 3.5-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent.
Precipitation increased measured pesticide concentrations xxx-fold. 
The highest concentrations were found during summer.
%%%! Add these numbers also to results section!
Risk thresholds were exceeded in 27 percent of small streams, with neonicotinoid and other insecticides exhibiting highest frequencies of exceedance. 
We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events in summer.

@@

Two recent studies indicate that pesticides might threaten freshwater biodiversity in the European union.
\citet{malaj_organic_2014} analyzed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that almost half of European water bodies are at risk from pesticides.
\citet{stehle_pesticide_2015} compiled 1,566 measured concentrations of 23 insecticides in the EU from scientific publications. 
They found that many of these measurements exceed regulatory acceptable concentrations (RAC).
However, these studies reflect only a small amount of potentially available data (173 sites in predominantly mid-sized and large rivers in \citet{malaj_organic_2014} and 138 measurements in \citet{stehle_pesticide_2015}), and it is unclear how representative they are for Germany. %175 estimated from digitized figure C.3 in Malaj 2014, Stehle: Table 2.
Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality,
which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs. 
Despite these data providing the opportunity to study pesticide risks and other research questions on a large scale with high spatial density, to date these data have not been compiled and related analyses are lacking.

Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
% needs definition of SWB
However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004,liess_determination_1999}.
Indeed, meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
Despite their ecological relevance and potentially higher pesticide exposure, a recent analysis of pesticide studies showed that a disproportionally small fraction of studies were conducted in SWB, and these were largely limited to a few sites \citep{lorenz_specifics_2016}. 
Consequently, knowledge on the pesticide pollution of SWB on larger scales is scant.

@@

%% -------------------------------------------------------------------------
\section{Methods}
\subsection{Data compilation}
We requested pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names). 
Additionally, we used data available from previous projects with bigger catchment sizes. 
We homogenized and unified all data provided by the federal states into a database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.

We identified chemical samples taken during heavy rainfall events by spatio-temporal intersection of sampling events with gridded daily precipitation data available from the German Weather service (DWD).
%%% add resolution 
This data spatially interpolates daily precipitation values from local weather stations \citep{rauthe_central_2013}. 
We performed the intersection for the actual sampling date and the day before, to identify rain events during and up to 48 hours before sampling. 
%%%! One question with such data is from which precipitation value increases in concentrations are detectable. Also interesting

\subsection{Characterization of catchments}
We compiled a total of 3,049 sampling sites with pesticide measurements.
We delineated catchments upstream for each of the sampling sites using a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
Catchment delineation was manually checked for accuracy by comparison with state stream networks.
Given that the delineated catchments were accurate only for 30\% of the sites, we compiled complementary catchment size data from authorities that provided catchments for an additional (60\% of sites). 
Ten percent of sites lacked catchment size data and were omitted from the operations outlined below.

For each derived catchment (either from DEM or drainage basins) we calculated the relative cover (in \%) with agricultural area based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}. 
Additionally, we used agricultural cover data provided by authorities (19\% of sites). 
For 78\% of the sites both, the proportion of agricultural land use and catchment size were available.
% see do_overview.R for numbers

@@

\subsection{Statistical analyses}
All data-processing and analyses were performed using R \citep{r_core_team_r:_2016}.
To display differences in the spectra of analyzed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with complete linkage hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
We modeled the number of RAC exceedances as:

\begin{align}
\begin{split}
  No(RQ > 1)_i \sim NB(\mu_i, \kappa) \\
  log(\mu_i)= \beta_0 + f_1(agri_i) + f_2(size_i) + log(n_i) \\
\end{split}
\end{align}

where $No(RQ > 1)_i$ is the observed number of RAC exceedances at site $i$. 
We modeled $No(RQ > 1)_i$ as resulting from a negative binomial distribution ($NB$) with mean $\mu_i$ and a quadratic mean-variance-relationship ($Var(No(RQ > 1)_i) = \mu_i + \frac{\mu_i^2}{\kappa}$).
The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of the number of RAC exceedances. 
$f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006} and $\beta_0$ is the intercept.
The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modeling the rate of exceedances. 
We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
GAMs were fitted using the mgcv package \citep{wood_fast_2011}.

To assess the influence of precipitation and seasonality we modeled the RQ of individual compounds as response variable.
RQ and concentrations show a skewed distribution with an excess of zeros (no pesticides detected and quantified). 
Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution (Equation~\ref{eqn:eqn3}) \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}.
These two processes can be interpreted as changes in the mean value of RQ (change in $\mu$) and changes in the probability of exceeding LOQ and showing any risk (change in $\nu$).

@@

$\nu_i$ denotes the probability of an observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values equal to or greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation of RQ.
We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($Q1-Q4$) as linear predictors for $\mu$ and $\nu$. 
We used appropriate link functions for $\mu$ and $\nu$ and assumed $\sigma$ to be constant. 
Equation~\ref{eqn:eqn4} summarises the deterministic part of the model.

@@

To account for temporal auto-correlation and differences between federal states we used $site$ nested within $state$ as random intercepts.
We implemented this model using the gamlss package \cite{stasinopoulos_generalized_2007}.

We fitted this model separately to each compound with a RAC, measured in least 1000 samples and with more than 5\% of values above LOQ (n = 24 compounds, see Supplemental Table~S3 for a list of compounds). 
To summarise the coefficients across the 24 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}. 
The results of individual compounds are provided in the Supplemental Table~S4 and Figure~S7.
The meta-analysis was performed using the metafor package \citep{Viechtbauer_2010}.

@@

The compiled dataset used in analysis comprised 2,918,604 pesticide measurements of 42,236 samples in 3,049 sampling sites.  %see do_overview.R for numbers.
These samples were taken via grab sampling.  
Composite samples of different durations that were taken in 33 sites were not considered in analysis, given a potential methodical bias and the low sample size.
We found large differences in the number of sampling sites between federal states and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1). 
The number of sampling sites per state ranged from 3 (Lower Saxonia, NI) to 1320 (North Rhine-Westphalia, NW). % Lower Saxony (NI) provided

@@

In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplemental Table~S2). 
Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
Most samples were taken in the months April till October, with fewer samples during winter (see Supplemental Figure~S3).
Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
We found substantial differences in the spectra of analyzed pesticides between federal states (Figure~\ref{fig:fig2}).
The number of different pesticides per state ranged from 57 (Saarland, SL) to 278 (Rhineland-Palatinate, RP). 
Hierarchical clustering revealed three groups (see also Supplemental Figure~ S4):

\begin{enumerate}[i)]
	\item SL, ST and TH with a small (less than 100 compounds) compound spectra.
	\item RP and NI with a big (278 and 226 compounds) and distinct compound spectra
  \item the other states showed a similar spectra of medium to big size (between 127 and 227 compounds)
\end{enumerate}

@@

The distribution of sampling sites across catchment sizes indicated a disproportionally low number of sites of catchments below $10~km^2$, with
most sampling sites having catchment sizes between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}). 
The proportion of agriculture in the catchments tended to decrease with increasing catchment size.
For example, only 3 sites had more than 75\% agriculture in catchments larger 100 $km^2$, whereas in catchments below 100 km2, 245 sites had more than 75\% agriculture in their catchment.

@@

\begin{figure}[ht]
  \includegraphics[width=0.95\textwidth]{figure4.pdf}
  \caption{Effect of percent agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances. Red line marks statistically significant changes. Dashed lines denote 95\% point-wise Confidence Intervals.
  }
  \label{fig:fig4}
\end{figure}

\subsection{Effect of precipitation on pesticide risk}
The spatio-temporal intersection revealed that most samples were taken during periods with low precipitation.
For example, only 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplemental Figure~S6).

$prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
An increase of precipitation before sampling ($prec_{-1}$) from 1mm to 10mm precipitation lead to a 6\% higher mean RQ in $Q2$.
The probability to exceed LOQ increases in $Q2$ from  12\% to 16\%.
Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$). 
This difference was less pronounced for the mean value of RQ (Figure~\ref{fig:fig5}, top).

The first quarter showed the lowest RQ and probability of exceeding LOQ.
Both increased during summer months and decreased towards the end of the year.
There was a probability of 4\% to exceed LOQ in $Q1$ and 10\% in $Q2$.
The differences were less pronounced for the value of RQ and with higher variation (Figure~\ref{fig:fig5}, bottom).

@@

\subsection{Pesticides in small water bodies}
The dataset comprised 12,710 samples from 1,295 small water bodies.
In 346 (26.7\%) of these streams RQ were exceeded.
In total 1,173 samples (0.3\% of all and 5.6\% of samples with detects) showed RQ higher than 1 were observed.
Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ exceedances (Figure~\ref{fig:fig6}).
For Thiacloprid, Imidacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (29.5, 14.2 and 21.4 \% of all samples with detects).
The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 82\% of all samples were it was analysed, see also Supplemental Figure~S9).
Glyphosate was the compound with the highest detection rates, followed by Boscalid and Isoproturon. 
However, only the latter showed RQ exceedances (Figure~\ref{fig:fig6}).
In 44.8\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8).

@@

%% -------------------------------------------------------------------------
\section{Discussion}
\subsection{Overview on the compiled dataset}
The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive one available for Germany.
Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
The data compiled and analysed here for Germany is of similar quantity and quality.

% current problems in monitoring and possible solutions
Nevertheless, a nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states. 
For Thiacloprid, Imidaclorpid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
For these compounds a lowering of LOQ is essential for reliable assessment.
Moreover, a nation-wide assessment would benefit from a harmonized spectrum of analysed compounds between federal states.

@@

\subsection{Influence of agricultural land use and catchment size}
% agriculture
We found a strong influence of agriculture on the pollution of streams.
If there is more the 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase the likelihood of RAC exceedance in fully agricultural catchments (above 75 \% agriculture).
To our knowledge this is the first study investigating such thresholds of pesticide risk.
Previous studies examined thresholds for percent agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
\citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use  in low land streams.
Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\%.
Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological changes.

% size
We did not find a relationship between pesticide pollution and catchment size.
However, previous studies showed that SWB are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
This can be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams with catchments above $10~km^2$ and below $100~km^2$ (Figure~\ref{fig:fig3}, top).
For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.

\subsection{Effect of precipitation on pesticide risk}
Our results revealed that pesticide sampling for chemical monitoring in Germany is mainly performed when no precipitation occurs. 
Nevertheless, we found higher RQ if samples where taken after rainfall events. 
Samples taken at the day of a rainfall-event showed less risk, than samples taken one day after a rainfall-event.
This could be explained by a sampling just before the actual rainfall event.
Pesticide concentrations in agricultural SWB generally show short term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after a rainfall-event might miss peak concentrations.
The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for RQ.
This could be attributed to a higher variability of absolute concentrations.
Overall, our results indicate that current pesticide monitoring relying on grab sampling, largely disconnected from precipitation events, underestimates pesticide risks.
Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for SWB \citep{lorenz_specifics_2016}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% 150-200 words
Small streams are important refugia for biodiversity.
In agricultural areas they may be at high risk from pesticide pollution. 
However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales. We used data from german governmental water quality monitoring to quantify the drivers of pesticide exposure and to assess pesticide risk in small streams on a large scale. The data set comprised of 2,918,604 measurements related to 42,236 samples in 3,049 sampling sites and of 484 pesticides. We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide exposure using new statistical modeling techniques that explicitly  consider the limit of quantification. 
\newline
\begin{itshape} die naechsten 2 Saetze sind noch schwach - hier muessen auch Zahlen hin - ausserdem sollten die Risiken klarer dargestellt werden inklusive der verantwortlichen Substanzen
\end{itshape}
\newline
Agriculture land use lead to an xxx-fold increase in xxx once the proportion of agriculture in a catchment exceeded yyy.
Precipitation during and before sampling increased measured pesticide concentrations xxx-fold. The highest concentrations were found during summer, though this varied with compound. Risk thresholds were exceeded in yyy percent of streams, with neonicotinoid and other insecticides exhibiting highest frequencies of exceedance. We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events in summer.
\newline
\begin{itshape} oder irgendsowas - auf jeden Fall irgendwas spektakulaerers als zu sagen dass pestizide ein risiko darstellen, dass ist nicht sonderlich neu
\end{itshape}

@@

Two recent studies indicate that pesticides might threaten freshwater biodiversity in the European union. 
\citet{malaj_organic_2014} analyzed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that almost half of European water bodies are at risk from pesticides.
\citet{stehle_pesticide_2015} compiled 1,566 measured concentrations of 23 insecticides in the EU from scientific publications. 
They found that many of these measurements exceed regulatory acceptable concentrations (RAC). 
However, these studies reflect only a small amount of potentially available data (173 sites in predominantly mid-sized and large rivers in \citet{malaj_organic_2014} and 138 measurements in \citet{stehle_pesticide_2015}), and it is unclear how representative they are for Germany. %173 see Schäfer et al. 2016, Stehle: Table 2. 
Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality, which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs. 
Despite these data providing the opportunity to study pesticide risks and other research questions on a large scale with high spatial density, to date these data have not been compiled and related analyses are lacking.

Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}
, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004,liess_determination_1999}.
Indeed, meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
Despite their ecological relevance and potentially higher pesticide exposure, a recent analysis of pesticide studies showed that a disproportionally small fraction of studies were conducted in SWB, and these were largely limited to a few sites \citep{lorenz_specifics_2016}. Consequently, knowledge on the pesticide pollution of SWB on larger scales is scant.

@@

%% -------------------------------------------------------------------------
\section{Methods}
\subsection{Data compilation}
We compiled pesticide monitoring data from sampling sites in streams with catchment sizes $\mathrm{< 150km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).

We homogenized and unified all data provided by the federal states into a database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.

We identified chemical samples taken during heavy rainfall events by spatio-temporal intersection of sampling events with gridded daily precipitation data available from the German Weather service (DWD).
This data spatially interpolates daily precipitation values from local weather stations \citep{rauthe_central_2013}. 
We performed the intersection for the actual sampling date and the day before, to identify rain events during and up to 48 hours before sampling.
%details required - what is a heavy rainfall event, resolution of data. One question with such data is from which precipitation value increases in concentrations are detectable. Also interesting and may be added as question in the introduction

\subsection{Characterization of catchments}
We compiled a total of 3,049 sampling sites with pesticide measurements.
We delineated catchments upstream for each of the sampling sites using a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
Catchment delineation was manually checked for accuracy by comparison with state stream networks.
Given that the delineated catchments were accurate only for 30\% of the sites,
%a reviewer might ask how accuracy was defined/cehcked
we compiled complementary catchment size data from authorities that provided catchments for an additional (60\% of sites). Ten \% of sites lacked catchment size data and were omitted from the operations outlined below.

For each catchment we calculated the relative cover (in \%) with agricultural area based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
%it may be worth mentioning, what agricultural land cover is
Additionally, we used agricultural cover data provided by authorities (19\% of sites). 
For 78\% of the sites both, the proportion of agricultural land use and catchment size were available.
% see do_overview.R for numbers
% this section is puzzling - if you have 90% catchments, why is it only 78% in the end? It is also unclear what is the ATKIS coverage and why additional state-wide??? cover data was used...

@@

\subsection{Statistical analyses}
All data-processing and analyses were performed using R \citep{r_core_team_r:_2016}.
To display differences in the spectra of analyzed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with complete linkage hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
We modeled the number of RAC exceedances (RQ \textgreater 1) as:

\begin{align}
\begin{split}
  No_i \sim NB(\mu_i, \kappa) \\
  % E(No_i) = \mu_i~and~Var(No_i) = \mu_i + \frac{\mu_i^2}{\kappa} \\
  log(\mu_i)= \beta_0 + f_1(agri_i) + f_2(size_i) + log(n_i) \\
\end{split}
\end{align}

where $No_i$ is the observed number of exceedances at site $i$.

We modeled $No_i$ as resulting from a negative binomial distribution ($NB$) with .....
The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of .... 
$f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}.
The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
The number of samples per site ($n_i$) was used as an offset to account for differences in sampling efforts (number of samplings and of analysed compounds) at a site and is equivalent to modeling the rate of exceedances.

We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
GAMs were fitted using the mgcv package \citep{wood_fast_2011}.

While agricultural land use and catchment size vary only between sites, this is not the case for precipitation which changes also with time.
Therefore, we modeled the effects of precipitation in a separate model.

RQ and concentrations show a skewed distribution with an excess of zeros (no pesticides detected and quantified). 
Therefore, we modeled these as two processes

(one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}:

@@

$\nu_i$ denotes the probability of an observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values equal to or greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation.
%weiß nicht, ob der normale leser kapiert, dass das mean und sd von RQ sind...
We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($Q1-Q4$) as linear predictors for $\mu$ and $\nu$. 
We used appropriate link functions for $\mu$ and $\nu$ and assumed $\sigma$ to be constant. 
Equation~\ref{eqn:eqn4} summarises the deterministic part of the model.

@@

To account for temporal auto-correlation and differences between federal states we used $site$ nested within $state$ as random intercepts.
Changes on $\mu$ can be interpreted as changes in the mean value of RQ, whereas changes in $\nu$ can be interpreted as changes in the probability of exceeding LOQ and showing any risk.

We implemented this model using the gamlss package \cite{stasinopoulos_generalized_2007}.

We fitted this model to each compound with a RAC, measured in least 1000 samples and with more than 5\% of values above LOQ (n = 24 compounds, see Supplemental Table~S3 for a list of compounds). 
To summarise the coefficients across the modeled compounds we used a random effect meta-analysis \citep{harrison_getting_2011}.
%müsste man vielleihct noch 2-3 sätze anfügen, was da genau gemacht wurde

@@

The compiled dataset used in analysis comprised 2,918,604 pesticide measurements of 42,236 samples in 3,049 sampling sites.  %see do_overview.R for numbers.
These samples were taken via grab sampling.  % see clean.R for numbers
Composite samples of different durations that were taken in 33 sites were not considered in analysis, given a potential methodical bias and the low sample size.
We found large differences in the number of sampling sites between federal states and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1).

@@

In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplemental Table~S2). 
Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
Most samples were taken in the months April till October, with fewer samples during winter (see Supplemental Figure~S3).
Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
We found substantial differences in the spectra of analyzed compounds between federal states (Figure~\ref{fig:fig2}).
Hierarchical clustering revealed three groups (see also Supplemental Figure~ S4):

\begin{enumerate}[i)]
	\item with less than 100 compounds (SL, ST and TH)
	\item with medium sized spectra
	\item with a big and distinct spectrum (RP and NI)
\end{enumerate}

@@

The distribution of sampling sites across catchment sizes indicated a disproportionally low number of sites in catchments below $10~km^2$, with most sampling sites in catchments between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}).
% vielleicht müsste man irgendwo nochmal schreiben, dass die zahl der kleinen gewässer zunimmt und man eben diese abnahme sieht in der verteilung
The proportion of agriculture in the catchments tended to decrease with increasing catchment size.
% put a statement here like: For example, only x% of sites had more than 75% agriculture in catchments larger 100 km2, whereas in catchments below 100 km2, yyy% of sites had more than 75% agriculture in their catchment.

@@

\begin{figure}[ht]
  \includegraphics[width=0.95\textwidth]{figure4.pdf}
  \caption{Effect of \% agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances. Red line marks statistically significant changes. Dashed lines denote 95\% point-wise Confidence Intervals.
  }
  \label{fig:fig4}
\end{figure}

\subsection{Effect of precipitation on pesticide exposure}
The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplemental Figure~S6).
$prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$). 
This difference was less pronounced for the mean value of RQ (Figure~\ref{fig:fig5}, top).

The first quarter showed the lowest RQ and probability of exceeding LOQ.
Both increased during summer months and decreased towards the end of the year.
The differences were less pronounced for the mean value RQ and with higher variation (Figure~\ref{fig:fig5}, bottom).

@@

\subsection{Pesticides in small water bodies}

The dataset comprised 12,710 samples from 1,295 small water bodies.
In 1,173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher than 1 were observed.
Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ exceedances (Figure~\ref{fig:fig6}).
For Thiacloprid, Imidacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (29.5, 14.2 and 21.4 \% of all samples with detects).
The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 82\% of all samples were it was analysed, see also Supplemental Figure~S9).
Glyphosate was the compound with the highest detection rates, followed by Boscalid and Isoproturon. 
However, only the latter showed RQ exceedances (Figure~\ref{fig:fig6}).
In 44.8\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8). 
% und man könnte sich fragen, ob es da unterschiede zwischen den bundesländern gab

@@

%% -------------------------------------------------------------------------
\section{Discussion}
\subsection{Overview on the compiled dataset}
The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive one available for Germany.
Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).

The data compiled and analysed here for Germany is of similar quantity and quality.

% current problems in monitoring and possible solutions
Nevertheless, a nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
% was wiederum die oben aufgeworfene frage verstärkt, ob es bei den ergebnissen unterschiede zwischen den staaten gab...
For Thiacloprid, Imidaclorpid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
For these compounds a lowering of LOQ is essential for reliable assessment.
Moreover, a nation-wide assessment benefit from a harmonized spectrum of analysed compounds between federal states.
% ok, aber dann musst du auch irgendwo positiver formulieren, worauf sich die ergebnisse von uns jetzt beziehen - schon mal nicht auf Deutschland...

@@

\subsection{Influence of agricultural land use and catchment size}
% agriculture
We found a strong influence of agriculture on the pollution of streams.
If there is more the 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase in the likelihood of RAC exceedance in fully agricultural catchments (above 75 \% agriculture).

To our knowledge this is the first study investigating such thresholds of pesticide risk.
Previous studies examined thresholds for \%agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
\citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use  in low land streams. 
Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\% agricultural land use for ....
Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological changes.

% size
We did not find a relationship between pesticide pollution and catchment size.
However, previous studies showed that SWB are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
This can be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams with catchments below \textless $100~km^2$ (Figure~\ref{fig:fig3}, top).
For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
Another reason might be the unequal distribution of catchment sizes, with fewer sites below $10~km^2$ and above $100~km^2$ catchment size (Figure~\ref{fig:fig3}, top).
